\documentclass{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Packages

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Definitions

\newtheorem{theorem}{Theorem}[subsection]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem*{remark}{Remark}
\newtheorem{definition}{Definition}[subsection]
\newtheorem{fact}{Fact}[subsection]
\newtheorem*{idea}{Idea}

\DeclareMathOperator{\spn}{Span}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator{\SU}{SU}

\newcommand{\bam}[1]{\textbf{#1}}
\newcommand{\mf}[1]{\mathfrak{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\comm}[2][]{\left[ #1, #2 \right]}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\set}[1]{\lbrace #1 \rbrace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Preamble

\title{Symmetries, Fields, and Particles Revision Notes}
\author{Linden Disney-Hogg}
\date{December 2018}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\maketitle
\tableofcontents

\section{Introduction}
A brief overview of some key ideas, concepts, and facts that I find useful in revising SFP. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lie Groups}
\subsection{Basics}
\begin{definition}[Lie Group]
A \bam{Lie group} is a group that has a smooth manifold structure such that the group operations are smooth functions on the manifold. 
\end{definition}

\begin{idea}
Lie groups are introduced here to give a foundation to the idea of a continuous symmetry group, which will act in some way. It is a consequence of the Peter-Weyl theorem that every compact Lie groups is isomorphic to a subgroup of $GL(n,\mbb{C})$ for some $n$ so often one can restrict their thought process to matrix Lie groups. 
\end{idea}

\begin{idea}
When showing that something is a structure is a Lie group, typically the hardest step is showing that it is a manifold. One approach is to find an explicit parametrisation of the manifold, often a good approach if the group has very obvious parameters upon which it depends (e.g. the Heisenberg group). Alternatively, one can use the preimage theorem, useful if the group is defined by some constraints (e.g. the orthogonal group). 
\end{idea}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lie Algebras}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Basics}

\begin{definition}[Lie Algebra] 

A \bam{Lie algebra} is a vector space, equipped with a bracket that:
\begin{itemize}
    \item is bilinear
    \item is antisymmetric
    \item obeys the Jacobi identity
\end{itemize}
\end{definition}

\begin{idea}
Given a matrix Lie group $G$, it has a natural associated Lie algebra $\mf{g}=T_{I}G$ with bracket given by the matrix commutator. This will be the prototypical Lie algebra to think of. The Lie algebra corresponding to a Lie group is useful to work with as we can take a basis, which will commute under addition, rather than work with a difficult presentation of the group.
Using the exponential map, we can map back into the original Lie group. This is not generally injective or surjective, but map injectively onto the connected component containing the identity. 
\end{idea}

\begin{fact}
The dimension of the tangent space to a manifold is equal to the dimension of the manifold. Hence, a Lie group and its associated Lie algebra have the same dimension
\end{fact}

\begin{definition}[Structure Constants]
Given a basis $\lbrace T^a \rbrace$ of a Lie algebra, the \bam{structure constants} $f^{ab}_c$ are defined such that 
\[
\comm[T^a]{T^b}=f^{ab}_c T^c
\]
\end{definition}

\begin{definition}[Complexification and Real Form]
Given a  real Lie algebra $\mf{g}_\mathbb{R}\equiv\spn_{\mathbb{R}}\lbrace T^a : a=1, \dots, D \rbrace$ the \bam{complexification}
is $\mf{g}_\mathbb{C}\equiv\spn_{\mathbb{C}}\lbrace T^a : a=1, \dots, D \rbrace$. Conversely, given such a $\mf{g}_\mathbb{C}$, $\mf{g}_\mathbb{R}$ is called a \bam{real form}.  
\end{definition}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ideals and Simplicity}

\begin{definition}[ideal]
An \bam{ideal} is a subalgebra $\mf{h} \leq \mf{g}$ s.t.
\[
\forall X\in\mf{g}, \; \forall Y\in\mf{h} \quad  \comm[X]{Y}\in\mf{h}
\]
\end{definition}

\begin{definition}[Derived Algebra]
The \bam{derived algebra} of a Lie algebra $\mf{g}$ is 
\[
i(\mf{g})=\set{\comm[X]{Y} : X,Y\in\mf{g}}
\]
\end{definition}

\begin{definition}[Centre]
The \bam{centre} of a Lie algebra $\mf{g}$ is 
\[
\zeta(\mf{g})=\set{X\in\mf{g} : \forall Y\in\mf{g} \; \comm[X]{Y}=0}
\]
\end{definition}

\begin{definition}[Simple Lie Algebra]
A Lie Algebra is \bam{simple} if it contains no non-trivial ideals.
\end{definition}

\begin{idea}
An ideal is to a Lie algebra what a normal subgroup is to a Lie group. Hence simple Lie algebras are the analogy of simple Lie groups.  
\end{idea}

\begin{definition}[Semi-Simple Lie Algebra]
A Lie Algebra is \bam{semi-simple} if it contains no non-trivial abelian ideals. 
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Adjoint Map}

\begin{definition}[Adjoint Map]
The \bam{adjoint map} of an element $X\in\mf{g}$ is $ad_{X}:\mf{g}\to\mf{g}$ defined such that for $Y\in\mf{g}$
\[
ad_{X}\left(Y\right)=\comm[X]{Y}
\]
\end{definition}

\begin{definition}[ad-diagonalisable]
An element $X\in\mf{g}$ is \bam{ad-diagonalisable} if the linear map $ad_X$ is diagonalisable. 
\end{definition}

\begin{definition}[Killing Form]
The \bam{Killing form} is a map $\kappa:\mf{g}\times\mf{g}\to\mbb{C}$ defined such that for $X,Y\in\mf{g}$
\[
\kappa\left(X,Y\right)=\tr\left(ad_X \circ ad_Y\right)
\]
It is symmetric and bilinear 
\end{definition}

\begin{theorem}
Let $\mf{g}$ be a Lie algebra and $\kappa$ the Killing form. Then
\[
\forall X,Y,Z\in\mf{g} \quad \kappa\left(\comm[Z]{X},Y\right)+\kappa\left(X,\comm[Z]{Y}\right)=0
\]
\end{theorem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cartan Subalgebras}

\begin{definition}[Cartan Subalgebra]
A \bam{Cartan subalgebra} $\mf{h}\leq\mf{g}$ is a maximal abelian subalgebra containing only ad-diagonalisable elements. It is typically written as $\mf{h}=\spn\lbrace H^i \rbrace$
\end{definition}

\begin{fact}
All Cartan subalgebras of a given Lie algebra have the same dimension. Though Cartan subgalgebras are not unique, they are all conjugate. 
\end{fact}

\begin{idea}
It is a fact that Cartan subalgebras correspond to maximal toral subgroups of the corresponding Lie group. This gives a route to visualisation of what the subalgebra might be. A Cartan subalgebra is defined such that adjoint maps corresponding to the basis $\lbrace H^i \rbrace$ are simultaneously diagonalisable. 
\end{idea}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{definition}[$\kappa^{ij}$]
Given a basis of a Cartan subalgebra $\lbrace H^i \rbrace$ and Killing form $\kappa$ we define 
\[
\kappa^{ij}=\kappa(H^i, H^j)
\]
\end{definition}

\begin{definition}[Compact Type]
A real Lie algebra $\mf{g}_\mathbb{R}$ is of \bam{compact type} if $\exists$ a basis in which $\kappa^{ij}=-K\delta^{ij}$ for some $K\in\mathbb{R}^{+}$.
\end{definition}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Representations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Basics}

\begin{definition}[Representations]
A \bam{representation} of a Lie group $G$ is a smooth group homomorphism 
\[
D:G\to GL\left(V\right)
\]
where $V$ is a $n$-dimensional vector space called the \bam{representation space}.
In analogy a representation of a Lie algebra is a Lie algebra homomorphism 
\[
d:\mf{g} \to \mf{gl}\left(V\right)
\]
\end{definition}

\begin{definition}[Dimension]
The \bam{dimension} of a representation is the dimension of the representation space.
\end{definition}

\begin{definition}[Trivial Representation] 
The \bam{trivial representation} is the unique one dimensional representation that sends every element to the identity. 
\end{definition}

\begin{definition}[Fundamental Representation]
For a matrix Lie group $G\leq GL(V)$ the \bam{fundamental representation} is the identity map $R=id_V$. Note $dim\left(R\right)=dim\left(V\right)$.
\end{definition}

\begin{definition}[Adjoint Representation]
Given a Lie algebra $\mf{g}$ the \bam{adjoint representation} is $d_{adj}:\mf{g}\to\mf{g}$, $d_{adj}\left(X\right)=ad_X $. Note $dim\left(d_{adj}\right)=dim\left(\mf{g}\right)$
\end{definition}

\begin{definition}[Faithful Representation]
A representation is \bam{faithful} if it is injective. 
\end{definition}

\begin{definition}[Isomorphic Representations]
Two representations $R_1$, $R_2$ of a Lie algebra $\mf{g}$ are \bam{isomorphic}, written $R_1 \cong R_2$ if $\exists S$ an invertible matrix such that 
\[
\forall X\in\mf{g} \quad R_2\left(X\right)=S R_1\left(X\right) S^{-1}
\]
\end{definition}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constructing Representations}

\begin{theorem}
A representation $D$ of a Lie group $G$ induces a representation on the the corresponding Lie algebra $\mf{g}$ as such: For $X\in\mf{g}$ let $g:(-\epsilon, \epsilon)\to G$ be a curve in $G$ defined such that $g(0)=e$ the identity of $G$ and $g'(0)=X$. This curve necessarily exists for some $\epsilon>0$. Then define $d$ by \[
d(X)\equiv\frac{d}{dt} D\left(g(t)\right) \vert_{t=0}
\]
Conversely, a representation of a Lie algebra $\mf{g}$ induces a representation in some neighbourhood of $e$ in $G$ given by 
\[
D(\exp{X})\equiv\exp{d(X)}
\]
\end{theorem}

\begin{definition}[Direct Sum Representation]
Given representations $R_1, R_2$ with representation spaces $V_1, V_2$ the \bam{direct sum representation} is $R_1 \oplus R_2$ with representation space $V_1 \oplus V_2$ acting by 
\[
(R_1 \oplus R_2)(X)(v_1 \oplus v_2)=(R_1(X)v_1)\oplus(R_2(X)v_2)
\]
\end{definition}

\begin{definition}[Tensor Product Representation]
Given representations $R_1, R_2$ with representation spaces $V_1, V_2$ the \bam{tensor product representation} is $R_1 \otimes R_2$ with representation space $V_1 \otimes V_2$ with
\[
(R_1 \otimes R_2)(X)=R_1(X)\otimes I_{V_2}+I_{V_1}\otimes R_2(X)
\]
\end{definition}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Reducibility}
\begin{definition}[Invariant Subspace]
Given a representation $R$ of Lie algebra $\mf{g}$ with representation space $V$, an \bam{invariant subspace} is $U\leq V$ such that
\[
\forall X\in\mf{g}, \; \forall u\in U \quad R\left(X\right)u\in U
\]
\end{definition}

\begin{definition}[Irreducible Representation]
A representation is \bam{irreducible} if it has no invariant subspaces. 
\end{definition}

\begin{definition}[Full Reducible]
A representation is \bam{fully reducible} if it can be expressed as the direct sum of irreducible representations. 
\end{definition}

\begin{fact}
If $R_i$ are finite dimensional irreducible representations of a simple Lie algebra for $i=1,\dots,m$ then
\[
R_1 \otimes R_2 \otimes \dots \otimes R_m
\]
is fully reducible. 
\end{fact}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weights and lattices}
\begin{definition}[Weights]
Given a representation $R$ of a Lie algebra with Cartan subalgebra $\mf{h}=\spn\lbrace H^i \rbrace$ the \bam{weights} of the representation are the eigenvalues of $R(H^i)$
\end{definition}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Useful Facts}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Miscellaneous}
\begin{theorem}
If $X$ is a complex square matrix then 
\[
\det{e^X}=e^{\tr{X}}
\]
\end{theorem}

\begin{theorem}[Baker-Campbell-Hausdorff Formula]
Let $\mf{g}$ be a Lie algebra, and $X,Y\in\mf{g}$. Then under sufficient existence conditions $e^X e^Y=e^Z$ where
\[
Z=X+Y+\frac{1}{2}\comm[X]{Y}+\frac{1}{12}\left(\comm[X]{\comm[X]{Y}}+\comm[Y]{\comm[Y]{X}}\right)+\dots
\]
\end{theorem}

\begin{corollary}[Lie Product Formula]
For $X,Y\in\mf{g}$, a Lie algebra
\[
e^{X+Y}=\lim_{N\to\infty}\left( e^{\frac{X}{N}} e^{\frac{Y}{N}} \right)^N
\]
\end{corollary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pauli Matrices}

\begin{definition}[Pauli Matrices]
The \bam{Pauli matrices} are

\begin{align*}
\sigma_1 &= \begin{pmatrix} 0 & 1 \\ 1 & 0\end{pmatrix}  \\
\sigma_2 &= \begin{pmatrix} 0 & -i \\ i & 0\end{pmatrix}  \\
\sigma_3 &= \begin{pmatrix} 1 & 0 \\ 0 & -1\end{pmatrix}  
\end{align*}

Note they are all Hermitian and traceless.
\end{definition}

\begin{fact}
\be
\sigma_i \sigma_j = \delta_{ij}I +i\epsilon_{ijk}\sigma_k
\ee
\end{fact}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SU(2)}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Basics}

\begin{definition}[U($n$)]
The \bam{unitary group} of dimension $n$ is 
\[
U\left(n\right)=\lbrace U\in GL\left(n,\mbb{C}\right) : U^\dagger U=I \rbrace
\]
It is path connected, and compact.
\end{definition}

\begin{definition}[SU($n$)]
The \bam{special unitary group} of dimension $n$ is
\[
SU\left(n\right)=\lbrace U\in U(n) : \det{U}=1 \rbrace
\]
\end{definition}
\noindent SU($n$) is a Lie group The corresponding Lie algebra is 
\[
\mf{su}\left(n\right)=\lbrace Z\in GL\left(n,\mbb{C}\right) : Z^\dagger+Z=0, \tr{Z}=0 \rbrace
\]
The real dimension is
\[
\underbrace{2\times\frac{1}{2}(n-1)n}_{\text{off diagonal elements}}+\underbrace{n}_{\text{diagonal elements}}-\underbrace{1}_{\text{trace constraint}}=n^2-1
\]
In the case $n=2$ we have 
\[
SU(2)=\lbrace   \begin{pmatrix} \alpha & -\overline{\beta} \\ \beta & \overline{\alpha} \end{pmatrix}  : \alpha,\beta\in\mathbb{C} , |\alpha|^2+|\beta|^2=1  \rbrace
\]
This can be expressed as, for $A\in SU(2)$
\[
A=a_0 I +i\bm{a}\cdot\bm{\sigma}
\]
where $\bm{a}=(a_1, a_2, a_3)$, $\bm{\sigma}=(\sigma_1, \sigma_2, \sigma_3)$, and $a_0^2+|\bm{a}|^2=1$. Hence $SU(2)\cong S^3$. In addition, by parametrising $SU(2)$ by the $a_i$, it can be seen that $\lbrace i\sigma_i\rbrace$ forms a basis of $\mf{su}(2)$. It is typical to normalise this basis to $\lbrace T^a=-\frac{1}{2} i\sigma_a \rbrace$. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Representations}

\begin{definition}[Cartan-Weyl basis of $\mf{su}_\mbb{C}(2)$]
$\lbrace H, E_\pm \rbrace$, where $H=\sigma_3$ and $E_\pm=\frac{1}{2}(\sigma_1\pm i\sigma_2 )$, is the \bam{Cartan-Weyl basis} for $\mf{su}_\mbb{C}(2)$. The elements satisfy the commutation relations 
\begin{align*}
\comm[H]{E_\pm} &= \pm 2E_\pm \\
\comm[E_+]{E_-} &= H
\end{align*}
\end{definition} 

\begin{fact}
$H$ spans a Cartan subalgebra of $\mf{su}_\mbb{C}(2)$, ad-diagonalisable w.r.t the Cartan Weyl basis, with eigenvalues $0, \pm 2$ respectively. 
\end{fact}

\begin{theorem}
Finite dimensional irreducible representations of $\mf{su}_\mbb{C}(2)$ with $R(H)$ diagonalisable are determined by the highest weight $\Lambda\in\mbb{N}_0$. Such a representation is called $R_\Lambda$. It has weight set 
\[
S_{R_\Lambda}=\lbrace \Lambda-2n : n=0,\dots,\Lambda \rbrace
\]
The eigenvectors of $R_\Lambda(H)$ are $v_{\Lambda-2n}$ defined such that 
\begin{align*}
R_\Lambda(H) v_\lambda &= \lambda v_\lambda \\
R_\Lambda(E_+) v_\Lambda &= 0 \\
 \left(R_\Lambda(E_-)\right)^n v_{\Lambda} &= v_{\Lambda-2n} \\
 R_\Lambda(E_+) v_{\Lambda-2n} &= r_n v_{\Lambda-2n}
\end{align*}
with 
\[
r_n=(\Lambda+1-n)n
\]
\end{theorem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cartan Classification}
\subsection{Root Geometry}

\begin{definition}[Roots]
The roots of a Lie Algebra are $\alpha\in\Phi$, vectors with components $\alpha^i$ eigenvalues of $ad_{H^i}$.
\end{definition}

\begin{definition}[$\alpha$-string through $\beta$]
The \bam{$\alpha$-string through $\beta$} is 
\[
S_{\alpha, \beta}=\lbrace \beta+\rho\alpha\in\Phi : \rho\in\mbb{Z} \rbrace
\]
\end{definition}

\begin{fact}
Every complex semi-simple Lie algebra of finite dimension has a real form of compact type.
\end{fact}

\end{document}
